---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# README

This repository contains the data and the analysis operations of the study titled "Pleasantness induction in music through consonance and dissonance: contrasting effect or tonal relation?" by Yuko Arthurs, Imre Lahdelma, and [Tuomas Eerola](https://tuomaseerola.github.io/).

## Load data

```{r, warning=FALSE,message=FALSE,echo=FALSE}
library(foreign)
library(tidyverse)
d <- foreign::read.spss("data/PleDisso1.sav")
df <- data.frame(d)

# convert into factors
df$id <- factor(df$id)
df$Age<-as.numeric(df$Age)
df$Gender<-factor(as.numeric(df$Gender),levels = c(1,2,3),labels = c("Male","Female","Other"))

df$Roughness<-factor(df$Roughness)
df$Familiarity<-factor(df$Familiarity)
df$NoDiaTone<-factor(df$NoDiaTone)
df$Musicianship<-factor(df$Musicianship)
df$OSMI<-factor(df$OSMI)

# Add stimulus type descriptions
s <- read.csv('data/stimulusTable.csv',header = TRUE)
#table(s$stimType)

stimulustype = c("HRHF_1","HRHF_2","HRLF_1","HRLF_2","LRHF_1","LRHF_2","LRLF_1","LRLF_2")
stimulusname = c("Major triad with added M2, root highest",
"Dominant 7th, 1st inversion",
"Minor-major 7th (no 5th, root doubled in oct)",
"Diminished + M7",
"Major triad + oct",
"Minor triad 1st inversion, 3rd doubled",
"Diminished 2nd inversion, 5th doubled",
"Power chord + M6d")

StimLabel<-c("Major+M2", "Dom7", "Min-maj7", "Dim+M7","Major+P8","Minor", "Dim", "Power+M6")
df$StimLabel <- factor(rep(StimLabel,7*50))

df$StimType<-NA
df$StimType<-rep(s$stimType,50)
df$StimTrans<-rep(s$stimTrans,50)

df$StimType<-factor(df$StimType,labels = stimulusname)
df$StimTrans<-factor(df$StimTrans)

# Make the levels explicit
df$Familiarity <- factor(df$Familiarity, levels = c(1,2),labels = c("Low","High"))
df$Roughness <- factor(df$Roughness, levels = c(1,2),labels = c("Low","High"))

df <- drop_na(df) # Delete missing data (1 participant)
df<-select(df,-NoDiaTone)

knitr::kable(head(df[,6:11]))

```

## Visualise

```{r plots,fig.width=10,fig.height=5,message=FALSE,warning=FALSE}
library(ggplot2)
library(tidyr)

S <- df %>%
  group_by(Roughness,Familiarity,StimLabel) %>%
  summarise(n=n(),m=mean(Rating,na.rm = TRUE),sd=sd(Rating,na.rm = TRUE)) %>%
  mutate(se=sd/sqrt(n),LCI=m+qnorm(0.025)*se,UCI=m+qnorm(0.975)*se) 

plsize <- 0.50 # errorbar line width
pd <- position_dodge(.5) # move them .05 to the left and right

S$RF<-paste(S$Roughness,S$Familiarity)

S$RF<-factor(S$RF,levels = c("High High","High Low","Low High","Low Low"), labels = c("HR-HF","HR-LF","LR-HF","LR-LF"))

g1 <- ggplot(S,aes(x=reorder(StimLabel,-m),y=m,shape=RF,color=RF))+
  geom_point(size=6,show.legend = TRUE)+
  geom_errorbar(aes(x=reorder(StimLabel,-m), ymin=LCI, ymax=UCI), width=0.2, linewidth = plsize,position = pd,show.legend = FALSE) + 
#  scale_fill_manual(values = c('red','blue'))+
#  scale_fill_manual(name="Roughness & Familiarity", values = c('red','blue'))+
  scale_color_manual(name="Roughness & Familiarity", values = c('black','black','grey50','grey50'))+
  scale_shape_manual(name="Roughness & Familiarity", values = c(16,17,21,24))+
  xlab('Chord Type')+
  ylab('Pleasantness (Mean ± 95% CI)')+
  scale_y_continuous(limits = c(4.5,6.3))+
  theme_bw()+
  theme(legend.justification=c(0,0), legend.position=c(0.77,0.5))+
  theme(text=element_text(size=14,  family="serif"))
print(g1)

```


## LMM Analysis of Two Factors

Here we run linear mixed model for Roughness and Familiarity and use participant and transposition as random effects.

```{r lmm1,warning=FALSE,message=FALSE}
library(lme4)
library(lmerTest)
library(emmeans)
library(effectsize)
```

## LMM Analysis of Two Fixed Factors and Three Random Factors

```{r lmm2c,warning=FALSE,message=FALSE,results='asis'}
df2<-dplyr::filter(df,Gender!='Other')

m1 <- lmer(Rating ~ Roughness * Familiarity + (1|id) + (1|StimTrans) + (1|Musicianship), data=df2,REML=FALSE)
print(knitr::kable(anova(m1)))

```
# Show Effect sizes using the partial epsilon squared

```{r effectsize,warning=FALSE,message=FALSE}
F_to_epsilon2(9.1497, 1, 2578.9) # 0.004 
F_to_epsilon2(204.00, 1, 2578.9) # 0.07
F_to_epsilon2(40.83, 1, 2578.9) # 0.02 
```

### Display Random Effects

```{r lmm2ranef,warning=FALSE,message=FALSE}
# add testing random effects
print(ranova(m1))
```


## LMM Analysis of Two Factors with additional information

Add musical expertise and gender and age to the mix. Here I first discard the two participants who did not disclose their Gender (for simplicity, we can treat this differently).

```{r lmm2b,warning=FALSE,message=FALSE}
df2<-dplyr::filter(df,Gender!='Other')

m2 <- lmer(Rating ~ Roughness * Familiarity + Age + Gender + Musicianship + Roughness:Age + Roughness:Gender + Roughness:Musicianship + Familiarity:Age + Familiarity:Gender + Familiarity:Musicianship+ (1|id) + (1|StimTrans), data=df2)
print(anova(m2))
```

OK, no interactions between the factors and background factors.

### LMM Analysis of 8 Chords

Are the 8 eight stimuli different from each other?

```{r lmm3,results='asis', eval=TRUE, message=FALSE,warning=FALSE,cache=TRUE}
library(papaja)
m1 <- lmer(Rating ~ StimLabel  + (1|id) + (1|StimTrans), data=df)
s2 <- summary(m1,corr=FALSE)

emm1<- emmeans(m1, specs = pairwise ~ StimLabel, adjust = "bonferroni") # are chords different between all pairings of the tracks?

apa_two <- apa_print(emm1$contrasts)
apa_table(
  apa_two$table
  , caption = "A full table of contrasts."
)
```


In sum, 19/28 comparisons are significant after correction for multiple comparison.


#### Session information

R version and libraries.

```{r}
print(sessionInfo())
```
